---
title: "Marketing Assignment 2"
author: "Bo Gu, Ke Deng, Lingyi Zhao"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    theme: united
    df_print: paged
    highlight: textmate
    toc: true
    toc_float: true
editor_options:
  chunk_output_type: inline
always_allow_html: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Loading in the data for the two datasets to understand what they contain.

```{r}
#install.packages("conjoint")
library(conjoint)
#data(journey)
data(chocolate)
```

**Journey** is a dataset around travel. The variables used in the conjoint analysis will be purpose w/ 4 levels(cognitive, vacation, health, business), form w/ 2 levels (organized vs. own), season w/ 2 levels (summer vs. winter), and accomodation w/ 4 levels (1-2-3 star_hotel, 4-5 star_hotel, guesthouse, hostel). 

**Chocolate** is a dataset around chocolate. The variables used in the conjoint analysis will be kind w/ 4 levels (milk,walnut, delicaties, dark), price w/ 3 levels(low, average, and high), packing w/ 2 levels (paperback vs. hardback),weight w/ 3 levels (light, middle, heavy) and calorie w/ 2 levels (little vs. much) 

description of the data provided in data(chocolate):

**cpref** = Vector of preferences (ratings) only - length 1392 (87 respondents x each respondent shown 16 profiles=1392).
**cprefm** = Matrix of preferences (87 respondents and 16 profiles).
**cprof** = Matrix of profiles (5 attributes and 16 full profiles).
**clevn** = Character vector of names for the attributes’ levels.
**csimp** = Matrix of simulation profiles.


#### PART 1


Use journey or chocolate data from package conjoint. Perform the following analyses: 

- Partworth estimation for all individuals separately 
- Partworth estimation by combining all individual responses into an aggregate 
- Partworth estimation by using linear mixed models. In this case, you should treat each individual as a separate group for defining the random effects 
- Partworth estimation of the mixed effects models by using MCMChregress

```{r}
set.seed(225)

#Randomly selecting the holdout data for the test dataframe
holdout <- vector()

#Randomly assigning 1-16 to each of the survey respondents 16 profiles
for (x in 1:87){
  j <- sample(1:16)
  holdout <- c(holdout,j)}

# adding in a response id column to identify each respondent - each respondent has a preference rating for 16 profiles
respid<- vector()
for (x in 1:87){
  j <- rep(x,16)
  respid <- c(respid,j)}

#establishing an inital dataframe
cprof.full <- cprof

#adding in cprof an additional 86 times, so cprof is repeated 87 times in cprof.full  

for (x in 1:86){
  cprof.full <- rbind(cprof.full,cprof)}

#creating a dataframe that includes all respondents with their 16 profiles and preferences stacked on top of each other. 

full.data <- data.frame(cbind(cpref, holdout, cprof.full,respid))

#to randomly remove 2 profiles per respondent going to use the randomly generated holdout data and just take the profiles where the holdout is 3-16 for the train and the test is 1 and 2. 

train.data <- full.data[full.data$holdout>2,]
test.data <- full.data[full.data$holdout<3,]
```

```{r}
#Model 1 - Individual Model
coeff <- matrix(nrow=87, ncol=15,dimnames = list(c(1:87),c('Constant',t(clevn))))
pred<- numeric()

#looping through individual lm models and taking the coefficients into a separate dataframe

cprof.factor=data.frame(apply(cprof,2,factor))

des.mm= model.matrix(~kind + price + packing + weight + calorie, data = cprof.factor,
              contrasts.arg=list(kind=contr.sum,price=contr.sum, 
                                 packing=contr.sum,weight=contr.sum, calorie=contr.sum))

des.mm.full <- des.mm

for (x in 1:86){
  des.mm.full <- rbind(des.mm.full,des.mm)
  }

des.df.full <- data.frame(des.mm.full)

chocolate <- data.frame(cbind(cpref, holdout,des.df.full,respid))
rownames(chocolate) <- 1:nrow(chocolate)

cho.train.data <- chocolate[chocolate$holdout>2,]
cho.test.data <- chocolate[chocolate$holdout<3,]

for (i in 1:87){
  model = lm(pref~kind1+kind2+kind3+price1+price2+packing1+weight1+weight2+calorie1, data = cho.train.data[cho.train.data$respid == i,])

  yhat = predict(model,cho.test.data[cho.test.data$respid == i,])
  pred = c(pred,yhat)
  coeff[i,1:4]=model$coefficients[1:4]
  coeff[i,5]=-sum(model$coefficients[2:4])
  coeff[i,6:7]=model$coefficients[5:6]
  coeff[i,8]=-sum(model$coefficients[5:6])
  coeff[i,9]=model$coefficients[7]
  coeff[i,10]=-model$coefficients[7]
  coeff[i,11:12]=model$coefficients[8:9]
  coeff[i,13]=-sum(model$coefficients[8:9])
  coeff[i,14]=model$coefficients[10]
  coeff[i,15]=-model$coefficients[10]
  }

# Write your code to run the individual lm models

mse_part1 <- mean(test.data$pref-pred)^2
mse_part1

```


```{r}
#Model 2 - Aggregate Model
#running full lm on all respondents and profiles as this is an aggregate model
#all respondents have the same coefficients

model_p2 = lm(pref~kind1+kind2+kind3+price1+price2+packing1+weight1+weight2+calorie1, data = cho.train.data)

summary(model_p2)

model_p2.parthworth <- model_p2$coefficients

model_p2.parthworth['kind4']=-sum(model_p2.parthworth['kind1'],model_p2.parthworth['kind2'],model_p2.parthworth['kind3'])
model_p2.parthworth['price3']=-sum(model_p2.parthworth['price1'],model_p2.parthworth['price2'])
model_p2.parthworth['packing2']=-sum(model_p2.parthworth['packing1'])
model_p2.parthworth['weight3']=-sum(model_p2.parthworth['weight1'],model_p2.parthworth['weight2'])
model_p2.parthworth['calorie2']=-sum(model_p2.parthworth['calorie1'])

model_p2_pred <- predict(model_p2,cho.test.data)


mse_part2 <- mean(test.data$pref-model_p2_pred)^2
mse_part2
```




```{r}
#Model 3 - Mixed Models Approach
require(lme4)

model_p3 <- lmer(pref~kind1+kind2+kind3+price1+price2+packing1+weight1+weight2+calorie1 + 
       (1+kind1+kind2+kind3+price1+price2+packing1+weight1+weight2+calorie1|respid),  
     control=lmerControl(optCtrl = list(maxfun=100000)), data=cho.train.data)

model_p3.parthworth <- matrix(nrow=87, ncol=15,dimnames = list(c(1:87),c('Constant',t(clevn))))

model_p3.coef <- coef(model_p3)$respid

for (i in 1:87){
  model_p3.parthworth[i,1]=model_p3.coef$`(Intercept)`[i]
  model_p3.parthworth[i,2]=model_p3.coef$kind1[i]
  model_p3.parthworth[i,3]=model_p3.coef$kind2[i]
  model_p3.parthworth[i,4]=model_p3.coef$kind3[i]
  model_p3.parthworth[i,5]=-sum(model_p3.parthworth[i,2:4])
  model_p3.parthworth[i,6]=model_p3.coef$price1[i]
  model_p3.parthworth[i,7]=model_p3.coef$price2[i]
  model_p3.parthworth[i,8]=-sum(model_p3.parthworth[i,6:7])
  model_p3.parthworth[i,9]=model_p3.coef$packing1[i]
  model_p3.parthworth[i,10]=-model_p3.parthworth[i,9]
  model_p3.parthworth[i,11]=model_p3.coef$weight1[i]
  model_p3.parthworth[i,12]=model_p3.coef$weight2[i]
  model_p3.parthworth[i,13]=-sum(model_p3.parthworth[i,11:12])
  model_p3.parthworth[i,14]=model_p3.coef$calorie1[i]
  model_p3.parthworth[i,15]=-model_p3.parthworth[i,14]
}

model_p3_pred <- predict(model_p3, cho.test.data)

mse_part3 <- mean(test.data$pref-model_p3_pred)^2
mse_part3
```



```{r}
#Model 4 - MCMC

require(MCMCpack)

model_p4 <- MCMChregress (fixed = pref~kind1+kind2+kind3+price1+price2+packing1+weight1+weight2+calorie1,
                          random =  ~1+kind1+kind2+kind3+price1+price2+packing1+weight1+weight2+calorie1,
                          group="respid",r=10, R=10*diag(10),data=cho.train.data)

#finding the fixed effects by taking the mean of the 1,000 simulations
fixed.intercept = mean(model_p4$mcmc[1:1000,1])
fixed.kind1 = mean(model_p4$mcmc[1:1000,2])
fixed.kind2 = mean(model_p4$mcmc[1:1000,3])
fixed.kind3 = mean(model_p4$mcmc[1:1000,4])
fixed.price1 = mean(model_p4$mcmc[1:1000,5])
fixed.price2 = mean(model_p4$mcmc[1:1000,6])
fixed.packing1 = mean(model_p4$mcmc[1:1000,7])
fixed.weight1 = mean(model_p4$mcmc[1:1000,8])
fixed.weight2 = mean(model_p4$mcmc[1:1000,9])
fixed.calorie1 = mean(model_p4$mcmc[1:1000,10])

model_p4_coeff=data.frame()

for (i in 1:87)
{
  #creating the variable name for the random effects, which are unique to respondent
  
  v1= mean(model_p4$mcmc[,paste("b.(Intercept).",as.character(i),sep = '')])
  v2= model_p4$mcmc[,paste("b.kind1.",as.character(i),sep = '')]
  v3= model_p4$mcmc[,paste("b.kind2.",as.character(i),sep = '')]
  v4= model_p4$mcmc[,paste("b.kind3.",as.character(i),sep = '')]
  v5= model_p4$mcmc[,paste("b.price1.",as.character(i),sep = '')]
  v6= model_p4$mcmc[,paste("b.price2.",as.character(i),sep = '')]
  v7= model_p4$mcmc[,paste("b.packing1.",as.character(i),sep = '')]
  v8= model_p4$mcmc[,paste("b.weight1.",as.character(i),sep = '')]
  v9= model_p4$mcmc[,paste("b.weight2.",as.character(i),sep = '')]
  v10= model_p4$mcmc[,paste("b.calorie1.",as.character(i),sep = '')]
  
  #taking the mean of the individual variables to find the random effects and adding the fixed effects
  model_p4_coeff[i,1] = mean(v1)+ fixed.intercept
  model_p4_coeff[i,2] = mean(v2)+ fixed.kind1
  model_p4_coeff[i,3] = mean(v3)+ fixed.kind2
  model_p4_coeff[i,4] = mean(v4)+ fixed.kind3
  model_p4_coeff[i,5] = mean(v5)+ fixed.price1
  model_p4_coeff[i,6] = mean(v6)+ fixed.price2
  model_p4_coeff[i,7] = mean(v7)+ fixed.packing1
  model_p4_coeff[i,8] = mean(v8)+ fixed.weight1
  model_p4_coeff[i,9] = mean(v9)+ fixed.weight2
  model_p4_coeff[i,10] = mean(v10)+ fixed.calorie1
}

model_p4.parthworth <- matrix(nrow=87, ncol=15,dimnames = list(c(1:87),c('Constant',t(clevn))))


for (i in 1:87){
  model_p4.parthworth[i,1]=model_p4_coeff$V1[i]
  model_p4.parthworth[i,2]=model_p4_coeff$V2[i]
  model_p4.parthworth[i,3]=model_p4_coeff$V3[i]
  model_p4.parthworth[i,4]=model_p4_coeff$V4[i]
  model_p4.parthworth[i,5]=-sum(model_p4.parthworth[i,2:4])
  model_p4.parthworth[i,6]=model_p4_coeff$V5[i]
  model_p4.parthworth[i,7]=model_p4_coeff$V6[i]
  model_p4.parthworth[i,8]=-sum(model_p4.parthworth[i,6:7])
  model_p4.parthworth[i,9]=model_p4_coeff$V7[i]
  model_p4.parthworth[i,10]=-model_p4.parthworth[i,9]
  model_p4.parthworth[i,11]=model_p4_coeff$V8[i]
  model_p4.parthworth[i,12]=model_p4_coeff$V9[i]
  model_p4.parthworth[i,13]=-sum(model_p4.parthworth[i,11:12])
  model_p4.parthworth[i,14]=model_p4_coeff$V10[i]
  model_p4.parthworth[i,15]=-model_p4.parthworth[i,14]
}

#generating the predictions by looping through and multiplying by the coefficients
model_p4_coeff_test = rbind(model_p4_coeff[1,],model_p4_coeff[1,])

for( i in 2:87){
  model_p4_coeff_test = rbind(model_p4_coeff_test,model_p4_coeff[i,],model_p4_coeff[i,])
}

model_p4_pred <- numeric()

# How will you score the test profiles in MCMC ?

for (i in 1:174){
  model_p4_pred[i] =  model_p4_coeff_test[i,1]+ 
                      model_p4_coeff_test[i,2]*cho.test.data[i,4]+ 
                      model_p4_coeff_test[i,3]*cho.test.data[i,5]+ 
                      model_p4_coeff_test[i,4]*cho.test.data[i,6]+
                      model_p4_coeff_test[i,5]*cho.test.data[i,7]+
                      model_p4_coeff_test[i,6]*cho.test.data[i,8]+
                      model_p4_coeff_test[i,7]*cho.test.data[i,9]+
                      model_p4_coeff_test[i,8]*cho.test.data[i,10]+
                      model_p4_coeff_test[i,9]*cho.test.data[i,11]+
                      model_p4_coeff_test[i,10]*cho.test.data[i,12]
}

mse_part4 <- mean(test.data$pref-model_p4_pred)^2
mse_part4

```


### Part 2

#### Compare the performance of the 4 models

```{r}
data.frame(mse_part1,mse_part2,mse_part3,mse_part4)
```


Based on the four models, the model with the lowest mse is **Model 3 - Mixed Models Approach** 

### Part 3

Now we will cluster the part-worths of the **Model 3 - Mixed Models Approach**.

```{r}
# Choose the coefficient from the best model based on test data performance
partworths <- coef(model_p3)$respid

# Perform Clustering
k.max <- 6
data <- partworths
wss <- sapply(1:k.max, 
              function(k){kmeans(data, k, nstart=50,iter.max = 15 )$tot.withinss})
wss
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")
```
**Based on the elbow method, we choose k=3.** 

```{r}
kmm = kmeans(data,3,nstart = 50,iter.max = 15) 
kmm
```
**After the k-means clustering, there are 26 respondants in cluster 1, 34 respondants in cluster 2, 27 respondants in cluster 3.**



### Part 4

Build a simulator that takes as input 4 or 5 input concepts, and produces market share estimates. Function caLogit can be used from package conjoint.

```{r}
##First, we build a dataframe of the profiles we want to use as inputs (calling this dataframe 'csimp0').

csimp0 <- data.frame(kind=c(4,1,2,3,2), 
                     price=c(1,3,2,1,3),
                     packing=c(2,1,1,1,2),
                     weight=c(2,1,3,1,3),
                     calorie=c(1,2,1,1,2))

##Here we can see what the dataframe looks like.
print(csimp0)
```


```{r}
#Then we run caLogit using 'csimp0' as our matrix of simulation profiles, along with the cpref and cprof given in the chocolate data. These are the market share estimates.

caLogit(csimp0,cpref,cprof)
```

### Part 5

**Cluster 1**

```{r}

cluster1 = kmm$centers[1,]
cluster1['kind4']= -sum(cluster1['kind1'], cluster1['kind2'], cluster1['kind3'])
cluster1['price3']=-sum(cluster1['price1'],cluster1['price2'])
cluster1['packing2']=-sum(cluster1['packing1'])
cluster1['weight3']=-sum(cluster1['weight1'],cluster1['weight2'])
cluster1['calorie2']=-sum(cluster1['calorie1'])
order(cluster1)

```
```{r}
cluster1[order(-cluster1)]
```


```{r}
cluster1.kind.importance = max(cluster1['kind1'],cluster1['kind2'],cluster1['kind3'],cluster1['kind4'])-min(cluster1['kind1'],cluster1['kind2'],cluster1['kind3'],cluster1['kind4'])
cluster1.price.importance= max(cluster1['price1'],cluster1['price2'],cluster1['price3'])-min(cluster1['price1'],cluster1['price2'],cluster1['price3'])
cluster1.packing.importance= max(cluster1['packing1'],cluster1['packing2'])-min(cluster1['packing1'],cluster1['packing2'])
cluster1.weight.importance= max(cluster1['weight1'],cluster1['weight2'],cluster1['weight3'])-min(cluster1['weight1'],cluster1['weight2'],cluster1['weight3'])
cluster1.calorie.importance= max(cluster1['calorie1'],cluster1['calorie2'])-min(cluster1['calorie1'],cluster1['calorie2'])

cluster1.attribute.importance = cbind(cluster1.kind.importance, cluster1.packing.importance, cluster1.price.importance,cluster1.weight.importance,cluster1.calorie.importance)
cluster1.attribute.importance
```

**Importance Ranking: kind > price > calorie > weight> packing **

**Top 5 profile in cluster 1:**

- Profile 1: dark, high price, much calorie, middle weight, hardback chocolate

- Profile 2: dark, high price, much calorie, middle weight, paperback chocolate

- Profile 3: dark, high price, much calorie, low weight, hardback chocolate

- Profile 4: dark, high price, much calorie, low weight, paperback chocolate

- Profile 5: dark, high price, much calorie, high weight, hardback chocolate




**Cluster 2**

```{r}
kmm$centers[2,]
```
```{r}

cluster2 = kmm$centers[2,]
cluster2['kind4']= -sum(cluster2['kind1'], cluster2['kind2'], cluster2['kind3'])
cluster2['price3']=-sum(cluster2['price1'],cluster2['price2'])
cluster2['packing2']=-sum(cluster2['packing1'])
cluster2['weight3']=-sum(cluster2['weight1'],cluster2['weight2'])
cluster2['calorie2']=-sum(cluster2['calorie1'])
order(cluster2)

```
```{r}
cluster2[order(-cluster2)]
```

```{r}
cluster2.kind.importance = max(cluster2['kind1'],cluster2['kind2'],cluster2['kind3'],cluster2['kind4'])-min(cluster2['kind1'],cluster2['kind2'],cluster2['kind3'],cluster2['kind4'])
cluster2.price.importance= max(cluster2['price1'],cluster2['price2'],cluster2['price3'])-min(cluster2['price1'],cluster2['price2'],cluster2['price3'])
cluster2.packing.importance= max(cluster2['packing1'],cluster2['packing2'])-min(cluster2['packing1'],cluster2['packing2'])
cluster2.weight.importance= max(cluster2['weight1'],cluster2['weight2'],cluster2['weight3'])-min(cluster2['weight1'],cluster2['weight2'],cluster2['weight3'])
cluster2.calorie.importance= max(cluster2['calorie1'],cluster2['calorie2'])-min(cluster2['calorie1'],cluster2['calorie2'])

cluster2.attribute.importance = cbind(cluster2.kind.importance, cluster2.packing.importance, cluster2.price.importance,cluster2.weight.importance,cluster2.calorie.importance)
cluster2.attribute.importance
```

**Importance Ranking: kind  > calorie > price > weight> packing **

**Top 5 profile in cluster 2:**

- Profile 1: dark, much calorie, high price, middle weight, hardback chocolate

- Profile 2: dark, much calorie, high price,  middle weight, paperback chocolate

- Profile 3: dark, much calorie, high price,  low weight, hardback chocolate

- Profile 4: dark, much calorie, high price,  low weight, paperback chocolate

- Profile 5: dark, much calorie, high price,  high weight, hardback chocolate



**Cluster3**

```{r}

cluster3 = kmm$centers[3,]
cluster3['kind4']= -sum(cluster3['kind1'], cluster3['kind2'], cluster3['kind3'])
cluster3['price3']=-sum(cluster3['price1'],cluster3['price2'])
cluster3['packing2']=-sum(cluster3['packing1'])
cluster3['weight3']=-sum(cluster3['weight1'],cluster3['weight2'])
cluster3['calorie2']=-sum(cluster3['calorie1'])
order(cluster3)

```
```{r}
cluster3[order(-cluster3)]
```


```{r}
cluster3.kind.importance = max(cluster3['kind1'],cluster3['kind2'],cluster3['kind3'],cluster3['kind4'])-min(cluster3['kind1'],cluster3['kind2'],cluster3['kind3'],cluster3['kind4'])
cluster3.price.importance= max(cluster3['price1'],cluster3['price2'],cluster3['price3'])-min(cluster3['price1'],cluster3['price2'],cluster3['price3'])
cluster3.packing.importance= max(cluster3['packing1'],cluster3['packing2'])-min(cluster3['packing1'],cluster3['packing2'])
cluster3.weight.importance= max(cluster3['weight1'],cluster3['weight2'],cluster3['weight3'])-min(cluster3['weight1'],cluster3['weight2'],cluster3['weight3'])
cluster3.calorie.importance= max(cluster3['calorie1'],cluster3['calorie2'])-min(cluster3['calorie1'],cluster3['calorie2'])

cluster3.attribute.importance = cbind(cluster3.kind.importance, cluster3.packing.importance, cluster3.price.importance,cluster3.weight.importance,cluster3.calorie.importance)
cluster3.attribute.importance
```

**Importance Ranking: kind  > calorie > price > weight> packing **


**Top 5 profile in cluster 3:**

- Profile 1: delicaties, much calorie, high price, heavy weight, paperback chocolate

- Profile 2: delicaties, much calorie, high price, heavy weight, haedback chocolate

- Profile 3: delicaties, much calorie, high price, middle weight, paperback chocolate

- Profile 4: delicaties, much calorie, high price, middle weight, hardback chocolate

- Profile 5: delicaties, much calorie, high price, low weight, paperback chocolate


